Gaps in stellar streams as a result of globular cluster fly-bys

### 2024-12-10 paper/main.tex
trying to contextualize bovik and bovy

  The simulations presented in this paper suggest that, in the last 5~Gyr of evolution, Palomar~5's stream could have experienced multiple close encounters with other Galactic globular clusters, some of which are able to create gaps -- even a few kpc wide -- in its tails.

  Unfortunately, none of the predicted gaps is located in regions of the tails that have been observed, and this is especially true for the leading tail, which result shorter than the trailing tail in the observations \citep[see][for a possible explanation of this feature]{2017NatAs...1..633P}. \textbf{NOT TRUE. Good thing I checked.}

  Currently, the presence -- or not -- of gaps in the observed portion of Pal~5 tails is debated: \citet{2016ApJ...819....1I} found no statistically significant gap in Pal~5 tails, while \citet{2017MNRAS.470...60E}, analyzing the same data-set as \citet{2016ApJ...819....1I}, suggested the presence of a few dips and gaps in the tails, at angular distances between $2^\circ$ and $9^\circ$ from the cluster center \citep[see also][]{ 2020ApJ...889...70B}.

  We cannot reproduce any of these features with clusters flybys in our simulations, since all our gaps are found at much larger angular and spatial distance from the cluster itself.

>>>>

  The simulations presented in this paper suggest that, in the last 5~Gyr of evolution, Palomar~5's stream could have experienced multiple close encounters with other Galactic globular clusters, some of which are able to create gaps -- even a few kpc wide -- in its tails. Currently, the presence -- or not -- of gaps in the observed portion of Pal~5 tails is debated: \citet{2016ApJ...819....1I} found no statistically significant gap in Pal~5 tails, while \citet{2017MNRAS.470...60E}, analyzing the same data-set as \citet{2016ApJ...819....1I}, suggested the presence of a few dips and gaps in the tails, at angular distances between $2^\circ$ and $9^\circ$ from the cluster center \citep[see also][]{ 2020ApJ...889...70B}. While our simulations produced 73 gaps across 50 realizations, only about 22 are beyond the current length of the observed portion of the stream. This means on average, we obtain at least one gap from a globular cluster within the past 5~Gyr. However, we note that we made no atttempt of creating mock observations to compare a full detection processes of gaps by disentangling field stars from stream stars. This is interesting work that is worth pursuing yet beyond the current scope of this work. The number of gaps present here is certainly an upper limit of detectable gaps.
